---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prioritizedeforestationhotspots

<!-- badges: start -->
[![R-CMD-check](https://github.com/albhasan/prioritizedeforestationhotspots/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/albhasan/prioritizedeforestationhotspots/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This package introduces a method for prioritizing deforestation hotspots to 
support law enforcement actions in the Brazilian Amazon, as describen in the 
paper entitled: *Science-based planning can support law enforcement actions to 
curb deforestation in the Brazilian Amazon* by Mataveli et at. (2022) [1], 
available at https://doi.org/10.1111/conl.12908

This is the abstract of the aforementioned paper [1]:

While Brazil publicly committed to reduce deforestation in Amazonia at the 26th
Conference of the Parties (COP26), the Brazilian parliament is moving toward 
weakening environmental laws. Deforestation rates continue ascending, reaching 
in 2021 the highest value since 2006 (13,235 km2). To overcome this paradox, 
strategies to curb deforestation are mandatory. The current strategy, “Plano 
Amazônia 21/22,” prioritizes law enforcement actions to curb illegal 
deforestation in only 11 Amazonian municipalities. Here, we show that this 
prioritization is likely to be insufficient since these municipalities account 
for just 37% of the current deforestation rate. This strategy may also be 
undermined by the leakage of deforestation actions to unmonitored 
municipalities. Using a set of spatially explicit datasets integrated into a 
deforestation-prediction modeling approach, we propose a science-based 
alternative method for ranking deforestation hotspots to be prioritized by law 
enforcement actions. Our prioritization method accounts for more than 60% of the 
deforestation, detecting larger deforested areas in both private and public 
lands, while covering 27% less territory than “Plano Amazônia 21/22.” Optimizing 
the detection of priority areas for curbing deforestation, as proposed here, is 
the first step to reducing deforestation rates and comply with the Brazilian 
legal commitment of 3925 km2 year−1.

[1] Mataveli, G., de Oliveira, G., Chaves, M. E., Dalagnol, R., Wagner, F. H., 
    Ipia, A. H., ... & Aragão, L. E. (2022). Science‐based planning can support 
    law enforcement actions to curb deforestation in the Brazilian Amazon. 
    Conservation Letters, e12908.



## Installation

You can install the latest version of prioritizedeforestationhotspots from [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("albhasan/prioritizedeforestationhotspots")
```

## Examples


### Run the model

This basic example produces results similar to those shown in the 
aforementioned paper (note that estimating the model accuracy takes hours!):

```{r example, eval=FALSE}

library(sf)
library(prioritizedeforestationhotspots)

# Run the model
out_dir <- "~/Documents/prioritize_res"
# NOTE: This takes long to run!
estimate_accuracy(out_dir)
fit_model(out_dir)
results_to_shp(out_dir)

```


### Plot the results (without runing the model)

If you don't have time to run the model, you can just load the results already
stored in the package.  This example shows the deforestation hotspots in a map:

```{r map_results, warning=FALSE, message=FALSE}

library(prioritizedeforestationhotspots)
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)
library(tidyr)
library(tidyselect)
library(tools)

# Read the result data from the package.
priority_sf <- 
    system.file("extdata", "results", "priority_classes.shp", 
                package = "prioritizedeforestationhotspots") %>%
    read_sf()

# Format the data.
priority_tb <- 
    priority_sf %>%
    st_drop_geometry() %>%
    pivot_longer(cols = starts_with("pri"), 
                 names_prefix = "pri",
                 names_to = "ref_year", 
                 values_to = "priority") 

# Arrange data into a sf object.
priority_sf <- 
    priority_sf %>%
    select(id) %>%
    right_join(priority_tb,  
               by = "id", 
               multiple = "all") %>%
    mutate(priority = factor(priority, 
                             labels = c("High", "Average", "Low"),
                             ordered = TRUE))

# Plot. 
priority_sf %>%
    ggplot() +
    geom_sf(aes(fill = priority), lwd = 0) +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    theme(legend.title=element_blank()) +
    ggtitle("Priority areas (deforestation hotspots)") 


```

We also included the variables used for the model.  This example shows some of 
the data included in this package:


### Plot the model variables

```{r map_data}

library(prioritizedeforestationhotspots)
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)

def_sf <- 
    deforestation_grid %>%
    right_join(deforestation_data, by = "id")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = area_PA), lwd = 0) +
    scale_fill_gradient(name = "Area (km2)", 
                        trans = "log",
                        breaks = c(1, 10, 100, 600),
                        low = "green",
                        high = "red") +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Protected areas or indigenous lands in the grid")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = dist_hidro), lwd = 0) +
    scale_fill_gradient(name = "Distance (km)", 
                        trans = "log",
                        breaks = c(1, 10, 50, 150),
                        low = "green",
                        high = "red") +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Distance to the closest waterway")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = dist_road), lwd = 0) +
    scale_fill_gradient(name = "Distance (km)", 
                        trans = "log",
                        breaks = c(1, 5, 50),
                        low = "green",
                        high = "red") +
    ggtitle("Distance to the closest highway")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = dist_road_hidro), lwd = 0) +
    scale_fill_gradient(name = "Distance (km)",
                        trans = "log",
                        breaks = c(1, 5, 50),
                        low = "green",
                        high = "red") +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Distance to the closest highway or waterway")

def_sf %>%
    drop_na(def) %>%
    ggplot() +
    geom_sf(aes(fill = def), lwd = 0) +
    scale_fill_gradient(name = "Deforestation (km2)",
                        trans = "log",
                        breaks = c(0.1, 1, 10, 100),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year, ncol = 1) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Deforestation")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = def_1_ly), lwd = 0) +
    scale_fill_gradient(name = "Deforestation (km2)",
                        trans = "log",
                        breaks = c(0.01, 0.10, 1, 10, 100),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Deforestation in the year before")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = def_2_ly), lwd = 0) +
    scale_fill_gradient(name = "Deforestation (km2)",
                        trans = "log",
                        breaks = c(0.01, 0.10, 1, 10, 100),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Deforestation 2 years before")

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = def_4_ly), lwd = 0) +
    scale_fill_gradient(name = "Deforestation (km2)",
                        trans = "log",
                        breaks = c(0.01, 0.10, 1, 10, 100),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Deforestation 4 years before")

def_sf %>%
    mutate(dist_1_percent_ly = dist_1_percent_ly + 1) %>%
    ggplot() +
    geom_sf(aes(fill = dist_1_percent_ly), lwd = 0) +
    scale_fill_gradient(name = "Distance (km)",
                        trans = "log",
                        breaks = c(25, 50, 100, 250, 500, 900),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle(paste("Distance to the closest grid centroid with more than 1%",
                  "deforestation in year before"))

def_sf %>%
    mutate(dist_2_percent_ly = dist_2_percent_ly + 1) %>%
    ggplot() +
    geom_sf(aes(fill = dist_2_percent_ly), lwd = 0) +
    scale_fill_gradient(name = "Distance (km)",
                        trans = "log",
                        breaks = c(25, 50, 100, 250, 500, 1000),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle(paste("Distance to the closest grid centroid with more than 2%",
                  "deforestation in year before"))

def_sf %>%
    ggplot() +
    geom_sf(aes(fill = active_fires_ly), lwd = 0) +
    scale_fill_gradient(name = "Number of fires",
                        trans = "log",
                        breaks = c(1, 5, 50, 500),
                        low = "green",
                        high = "red") +
    facet_wrap(~ref_year) +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Number of active fires in the year before")

```
