---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prioritizedeforestationhotspots

<!-- badges: start -->
<!-- badges: end -->

This package introduces a method for prioritizing deforestation hotspots to support law enforcement actions in the Brazilian Amazon, as describen in the paper entitled: *Science-based planning can support law enforcement actions to curb deforestation in the Brazilian Amazon*.


## Installation

You can install the development version of prioritizedeforestationhotspots from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("albhasan/prioritizedeforestationhotspots")
```

## Example

This basic example produces results similar to those shown in the aforementioned paper:

```{r example}
library(prioritizedeforestationhotspots)

out_dir <- "~/Documents/prioritize_res"

#estimate_accuracy(out_dir)
#fit_model(out_dir)

```

